On the Proof Theory of Property-Based Testing of Coinductive Specifications, or:\\ PBT to Infinity and beyond

### 2020-10-16 papers/ijcar2020/response-to-rebuttal.txt

>>  ----------------------- REVIEW 1 ---------------------

>>  ... My understanding is that they have a Coq tactic that takes
>>  a proof goal, exports to to an external mini-prover, which generates
>>  an FPC certificate. Then the FPC certificate is imported back to Coq
>>  via a deep embedding, and Coq uses a reflection theorem to turn it
>>  into an ordinary proof. But I am only guessing: this is not made
>>  clear.

A more accurate outline is: an external proof certificate is checked
via lambda Prolog (ELPI running within Coq).  If that check is
successful, lambda Prolog will build a proof structure that the Coq
kernel should be able to check (only the Coq kernel needs to be
trusted).  Part of the novelty here is the use of FPCs and logic
programming, which makes defining external proof structures, checking
such proofs, and generating Coq proofs all part of the same, modular
process.  Another novelty here is that Coq proofs are built entirely
within the ELPI plugin without using, say, Ltac or reflection.

>>  The example, a proof of "(A -> B) -> A -> B", is singularly
>>  unimpressive. Can the tool do something more complex? How long does it
>>  take?

Since this is an ongoing project, there are now more examples in the
GitHub repository.  Many other kinds of FPCs are described in 
papers [5,9,11].

>>  I don't understand what is in the repository. The READMEs don't
>>  provide any guidance.

The repository has been reorganized to make it much more clear.

>>  ----------------------- REVIEW 2 ---------------------

>>  - The presented example (example 1) that uses the system is
>>    trivial. It leaves me wondering how easy it is to apply the system
>>    to realistic examples.

There are more examples in the GitHub repository.

>>  * I am uncertain of the accuracy of the second sentence in the related
>>    work section. It states that this version of FPC-Coq only works for
>>    proving theorems in first order intuitionistic logic. Then continues
>>    to state that FPC-Coq can be used to prove the double negation of a
>>    formula for which there is a classical proof. Double negating a
>>    formula and finding an intuitionistic proof is only true for
>>    propositional logic and not first-order logic. First order negative
>>    translations are more involved (e.g. also double negating the body
>>    of each forall quantifier).

We meant to say "double negation translation" and not "the double
negation."  As this reviewer points out, many more negations need to
be inserted into first-order formulas for this scheme to work.  The
use of such double-negation translations in this focused setting has
been studied more carefully in reference [11] and in the technical
report https://hal.inria.fr/hal-01354298 (by Chihani, Ilik, Miller).

>>  - Section 2.1, could you please provide a short explanation why there
>>    are 101 introduction rules (and not 100)?

Since there are 100 formulas on the left and one on the right, there
are possibly 101 places to use an introduction rule.

>>  - Section 2.2: it may be helpful to explicitly state that decide and
>>    store are examples of experts and not necessarily part of every
>>    system.

We can be more clear about this: Decide and store are part of the
focusing mechanism and, hence, part of every FPC.  They are not,
however, part of the usual presentation of the sequent calculus.

### 2020-10-16 papers/ijcar2020/reviews-from-rebuttal.txt

IJCAR 2020 review response (submission [*NUMBER*])

Dear Dale Miller,

Thank you for your submission to IJCAR 2020. The IJCAR 2020
rebuttal period will be between March 9 and March 12 (AoE).

During this time, you will have access to the current state of
your reviews and have the opportunity to submit a response.
Please keep in mind the following during this
process:

* The response must focus on any factual errors in the reviews
and any questions posed by the reviewers. It must not provide
new research results or reformulate the presentation. Try to be
as concise and to the point as possible.

* The rebuttal period is an opportunity to react to the
reviews, but not a requirement to do so. Thus, if you feel the
reviews are accurate and the reviewers have not asked any
questions, then you do not have to respond.

* The reviews are as submitted by the PC members, without any
coordination between them. Thus, there may be inconsistencies.
Furthermore, these are not the final versions of the reviews.
The reviews can later be updated to take into account the
discussions at the program committee meeting, and we may find
it necessary to solicit other outside reviews after the
rebuttal period.

* The program committee will read your responses carefully and
take this information into account during the discussions. On
the other hand, the program committee will not directly respond
to your responses, either before the program committee meeting
or in the final versions of the reviews.

* Your response will be seen by all PC members who have access
to the discussion of your paper, so please try to be polite and
constructive.

The reviews on your paper are attached to this letter. To
submit your response you should log on the EasyChair Web page
for IJCAR 2020 and select your submission on the menu.

----------------------- REVIEW 1 ---------------------
SUBMISSION: 122
TITLE: FPC-Coq: Using ELPI to elaborate external proof evidence into Coq proofs (System Description)
AUTHORS: Roberto Blanco, Matteo Manighetti and Dale Miller

----------- Overall evaluation -----------
SCORE: -2 (reject)
----- TEXT:

This paper describes a connection between Coq and a lambda-Prolog that
allows one to generate proof certificates externally and verify them
in Coq. Even as a short paper, it is not suitable for publication in
its present form.

Four out of the six pages are devoted to describing the FPC proof
format, which has been described in detail in other publications. So
the "new" part is only two pages.

The paper is virtually unreadable. It is hard to understand what has
been done. My understanding is that they have a Coq tactic that takes
a proof goal, exports to to an external mini-prover, which generates
an FPC certificate. Then the FPC certificate is imported back to Coq
via a deep embedding, and Coq uses a reflection theorem to turn it
into an ordinary proof. But I am only guessing: this is not made
clear.

The example, a proof of "(A -> B) -> A -> B", is singularly
unimpressive. Can the tool do something more complex? How long does it
take?

The last page and a half drops down to a smaller font. Why? To fit in
extra text? That is cheating.

The paper is very sloppily written. There are very many typographical
errors, spelling errors, grammatical errors, sentences that don't make
sense, and a missing citation. I wonder whether the authors even
proofread the paper.

I don't understand what is in the repository. The READMEs don't
provide any guidance. There is no information as to how to install the
system and its dependencies, or how to test anything. There are only a
couple of hundred lines of Coq code, and it is not clear what they
do. There are number of commits made after the submission deadline.

----------------------- REVIEW 2 ---------------------
SUBMISSION: 122
TITLE: FPC-Coq: Using ELPI to elaborate external proof evidence into Coq proofs (System Description)
AUTHORS: Roberto Blanco, Matteo Manighetti and Dale Miller

----------- Overall evaluation -----------
SCORE: 2 (accept)
----- TEXT:
Summary:

This paper describes system FPC-Coq that combines two systems: the
foundational proof certificate framework and the Coq proof
assistant. This allows for using Coq to check partial proofs that are
missing some detailed information and not formed by explicit
lambda-terms. This paper extends the FPC framework to allow for proof
focussing and for augmenting proofs with certificates (that may be
used to fill in the missing information in the proof term).

Feedback and Questions:

+ The paper is well-written and easy to follow

+ The tool developed is exciting and seems potentially useful in proof
exchange between theorem provers.

- The presented example (example 1) that uses the system is
  trivial. It leaves me wondering how easy it is to apply the system
  to realistic examples.

* I am uncertain of the accuracy of the second sentence in the related
  work section. It states that this version of FPC-Coq only works for
  proving theorems in first order intuitionistic logic. Then continues
  to state that FPC-Coq can be used to prove the double negation of a
  formula for which there is a classical proof. Double negating a
  formula and finding an intuitionistic proof is only true for
  propositional logic and not first-order logic. First order negative
  translations are more involved (e.g. also double negating the body
  of each forall quantifier). Am I misunderstanding this paragraph, is
  this an error, or is there a justification for this that I am
  missing?

Further clarifications and typos:

- Section 2.1, could you please provide a short explanation why there
  are 101 introduction rules (and not 100)? 
- Section 2.1, it would be nice at the end of the section to go back
  to the example at the beginning and give the number of choices for
  this example using the focused proof system (with a short
  explanation). 
- Section 2.2: it may be helpful to explicitly state that decide and
  store are examples of experts and not necessarily part of every
  system. 
- There is a missing reference in the related work section [1,2, ?].

----------------------- REVIEW 3 ---------------------
SUBMISSION: 122
TITLE: FPC-Coq: Using ELPI to elaborate external proof evidence into Coq proofs (System Description)
AUTHORS: Roberto Blanco, Matteo Manighetti and Dale Miller

----------- Overall evaluation -----------
SCORE: 2 (accept)
----- TEXT:

The FPC-Coq system is built using Coq-ELPI that is a new Coq plugin
based on the ELPI implementation of λProlog. The foundational proof
certificate (FPC) framework is a system capable of checking and
elaborating a wide range of proof evidence and in FPC-Coq, it is
implemented using the logic programming language λProlog. The authors
propose the recently released Coq-ELPI plugin or Coq in which the ELPI
implementation of λProlog can send information to the Coq kernel.

Then it is possible to provide an additional procedure (by going
around Coq’s tactic language) for building and submitting formal
proofs to be checked by the Coq kernel. Thus, proof certificates
coming from external provers or from users can be converted to proofs
appropriate for the Coq kernel. Then, in this context, it is an
significant contribution to the FPC framework.

Section 2 presents the FPC framework with a focus on two problems with
attempting to use sequent proofs for theorem proving and proof
checking. The first is the lack of focus that is addressed in the
paper with a subsection dedicated to focused proof systems. The second
one is the way to determine from where missing information comes and
then another subsection considers how to augment the proof system with
certificates.  This section is well-written with enough details to
understand the questions and the approach followed by the authors and
its specificity. They are illustrated through some examples of FPC
specifications

Section 3 presents FPC-Coq and the way proposed to make Coq interact
with proof certificates. Moreover the Elpi extension for Coq allows
some facilities for building tactics written in λProlog.

This paper is a system description that is well written with enough
details about the design and the implementation and also with a
logical account of some steps.  It is a part of a more global project
that consists of using λProlog within a type theory setting, in
general, and within Coq, in particular.  The conversion of external
proof certificates to proofs appropriate for the Coq kernel is an
interesting and significant result and its implementation through
FPC-Coq will have an impact on some further formal proof developments.
Then I recommend to accept this system description.

Other comments:

p5. analisis -> analysis

P7. In related work paragraph a reference is missing [1,2,?]

------------------------------------------------------

Best wishes,

Nicolas Peltier
Viorica Sofronie-Stokkermans

### 2020-10-16 papers/coLP/types20/t20.tex
Added another paper.

\RequirePackage[utf8]{inputenc}
\RequirePackage[T1]{fontenc}
\documentclass[a4paper]{easychair}
\usepackage{url,xcolor}
\usepackage{proof}
\usepackage{listings}
\usepackage{letltxmacro}

\newcommand{\instan}[1]{\hbox{\sl grnd}~(#1)}
\newcommand{\Pscr}{{\mathcal P}}

\newcommand{\ok}{\checkmark}
% DM The use of pifont and txfonts broke typesetting for me: equality
% signs never appearred and some parentheses did not appear either.
%\newcommand{\noc}{\ding{56}}
\newcommand{\noc}{\ddag}
\newcommand{\lP}{$\lambda$Prolog\xspace}
\newcommand{\nat }{\hbox{\sl nat}\xspace}
\newcommand{\plus}{\hbox{\sl app}\xspace}
\newcommand{\lst}{\hbox{\sl lst}\xspace}
\newcommand{\sort}{\hbox{\sl sort}}
\newcommand{\rev}{\hbox{\sl rev}}
\newcommand{\ordered}{\hbox{\sl ordered}}

%\newcommand{\atac}{A2Tac\xspace}
\newcommand{\atac}{ACheck\xspace}
%\newcommand{\lra}{\mathrel{\longrightarrow}}
\newcommand{\lra}{\mathrel{\vdash}}
\newcommand{\seq}[2]{#1\lra #2}

\newcommand{\true}{tt}

%%%%%%%%%%%%%%%% LJF
\newcommand{\truen}{t^-}
\newcommand{\truep}{t^+}
\newcommand{\falsen}{f^-}
\newcommand{\falsep}{f^+}
\newcommand{\wedgep}{\wedge^{\!+}}
\newcommand{\wedgen}{\wedge^{\!-}}
\newcommand{\veep}{\vee^{\!+}}
\newcommand{\veen}{\vee^{\!-}}
\newcommand{\with}{\&}

\newcommand{\Nscr}{{\cal N}}
\newcommand{\Rscr}{{\cal R}}                                   % Used for an ambiguous rhs
%\newcommand{\Rscr}{\Delta_1\Downarrow\Delta_2}                % Used for an ambiguous rhs
\newcommand{\jUnf    }[4]{#1\mathbin\Uparrow#2\vdash#3\mathbin\Uparrow #4} % unfocused sequent
\newcommand{\jUnfG   }[2]{\jUnf{\Gamma}{#1}{#2}{{}}}           % unf sequ with \Gamma
\newcommand{\jUnfamb }[3]{#1\mathbin\Uparrow#2\vdash#3 \Rscr}  % unfocused sequent
\newcommand{\jUnfGamb}[1]{\Gamma\mathbin\Uparrow#1\vdash \Rscr}% unf sequ with \Gamma
\newcommand{\jLf     }[3]{#1\Downarrow#2\vdash#3}              % left focused sequent
\newcommand{\jLfG    }[1]{\jLf{\Gamma}{#1}{E}}                 % left foc seq with \Gamma 
\newcommand{\jRf     }[2]{#1\vdash #2\Downarrow}               % right focused sequent
\newcommand{\jRfG    }[1]{\jRf{\Gamma}{#1}}                    % right foc seq with \Gamma
%%%%%%%%%%%%%%%% 
\newcommand{\bxi}[1]{\blue{\Xi_{#1}}}
\newcommand{\bXi}[1]{\blue{\Xi_{#1} :\null}}

\newcommand{\andClerk}[3]{{\wedge_c}(#1,#2,#3)}
\newcommand{\falseClerk}[2]{f_c(#1,#2)}
\newcommand{\orClerk}[2]{{\vee_c}(#1,#2)}
\newcommand{\allClerk}[2]{\forall_c(#1,#2)}
\newcommand{\storeClerk}[3]{\hbox{\sl store}_c(#1,#2,#3)}

\newcommand{\trueExpert }[1]{{\true_e}(#1)}
\newcommand{\eqExpert }[1]{{=_e}(#1)}
\newcommand{\unfoldExpert}[2]{{\hbox{\sl unfold}_e}(#1,#2)}
\newcommand{\andExpert}[3]{{\wedge_e}(#1,#2,#3)}
\newcommand{\andExpertLJF}[6]{{\wedge_e}(#1,#2,#3,#4,#5,#6)}
\newcommand{\orExpert  }[3]{{\vee_e}(#1,#2,#3)}
\newcommand{\someExpert}[3]{\exists_e(#1,#2,#3)}
\newcommand{\initExpert}[2]{\hbox{\sl init}_e(#1,#2)}
\newcommand{\cutExpert}[4]{\hbox{\sl cut}_e(#1,#2,#3,#4)}
\newcommand{\decideExpert}[3]{\hbox{\sl decide}_e(#1,#2,#3)}
\newcommand{\releaseExpert}[2]{\hbox{\sl release}_e(#1,#2)}
%%%%%%%%%%%%%%%%

%
\newcommand{\step}{\longrightarrow}
\newcommand{\sstlc}{\textit{STLC}\xspace}
\newcommand{\vds}{\vdash_\Sigma}

\def\bnfas{\mathrel{::=}}
\def\bnfalt{\mid}

\def\lam{\lambda}
\def\Lam{\Lambda}
\def\arrow{\rightarrow}
\def\oftp{\mathord{:}}
\def\hastype{\mathrel{:}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Editorials
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\long\def\ednote#1{\footnote{[{\it #1\/}]}\message{ednote!}}
%\long\def\ednote#1{\begin{quote}[{\it #1\/}]\end{quote}\message{note!}}
\newenvironment{metanote}{\begin{quote}\message{note!}[\begingroup\it}%
                         {\endgroup]\end{quote}}
\long\def\ignore#1{}

\newcommand{\todo}[1]{\begin{metanote}TODO: #1\end{metanote}}
%
%\setlength{\textwidth}{13.2cm}
%\setlength{\textheight}{19.9cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%

\input{types20/listing-macros}  % Use to get better highlighing of code
\lstset{language=lprolog}
\lstset{language=abella}
\title{On the Proof Theory of Property-Based Testing of Coinductive Specifications, or:\\
 PBT to Infinity and beyond}
\authorrunning{Blanco, Miller and Momigliano}
\titlerunning{PBT to Infinity and
   beyond}
\author{Roberto Blanco\inst{1} \and Dale Miller\inst{2} \and Alberto Momigliano\inst{3}}

@@

\section{Introduction}
\label{sec:intro}

\begin{center}
  [For the moment taken from the TYPES abstract, to be revised]
\end{center}

@@

While PBT originated in the functional programming community, we have
given in a previous paper (\cite{Blanco0M19}) a reconstruction of some
of its features (operational semantics, different flavors of
generation, shrinking) in purely proof-theoretic terms employing the
framework of \emph{Foundational Proof Certificates}~\cite{chihani17jar}: the
latter, in its full generality, defines a range of proof structures
used in various theorem provers such as resolution refutations,
Herbrand disjuncts, tableaux, etc.
%
In the context of PBT, the proof theory setup is much simpler.

\begin{center}
  [etc\dots]
\end{center}

The contributions of this paper are
\begin{itemize}
\item Tighten the link of our previous reconstruction of PBT to the
  proof-theoretic approach of model checking in~\cite{heath19jar} by casting it in the setting of $\mu$MALL
\item Extend PBT beyond the Horn-nabla fragment to Harrop spec, first using only case analysis, then also co-induction
\item Experiment with a multi-level architecture, where FPC only takes
  care of generation and testing  is delegated to a system such as Bedwyr
\item Hopefully provide some examples (CCS, $\pi$ calculus (?), co-evaluation)
\item
  \begin{center}
[others \dots]    
  \end{center}

\end{itemize}

\section{The story so far}
\label{sec:tecap}

\begin{center}
  [Here we recap what we did in PPDP, in the uniform proof style, see fig~\ref{fig:augmented}: the connection with linear logic comes in the next section]
\end{center}

Consider an attempt to find counter-examples to a conjecture of the
form \(\forall x [(\tau(x)\wedge P(x)) \supset Q(x)]\) where $\tau$ is
a typing predicate and $P$ and $Q$ are two other predicates defined
using Horn clause specifications.
%
By negating this conjecture, we attempt to find a (focused) proof of 
\(\exists x [(\tau(x)\land P(x)) \land \neg Q(x)]\).
%
In the focused proof setting, the \emph{positive phase} (where
test cases are generated) is represented by \(\exists x\) and
\((\tau(x)\land P(x))\). 
%
That phase is followed by the \emph{negative phase} (where conjectured
counter-examples are tested) and is represented by \(\neg Q(x)\).
%
%% turns out to be a search for a proof of \(\exists x [(\tau(x)\land
%%   P(x)) \land \neg Q(x)]\) in a focused sequent calculus for
%% (basically) Horn logic, and pre-condition phase, while proving $Q$ can
%% be relegated to deterministic logic programming-like computation (the
%% \emph{negative} phase), interpreting ``$\neg$'' as
%% Negation-as-failure..
%
%

\begin{figure}
\newcommand{\XXi}{{\color{blue}{\Xi}}}
\[
\infer{\XXi\vdash G_1\wedge G_2}
      {\XXi_1\vdash G_1\qquad \XXi_2\vdash G_2\qquad \andExpert{\XXi}{\XXi_1}{\XXi_2}}
\qquad
\infer{\XXi\vdash \true}
      {\trueExpert{\XXi}}
\]
\vskip -6pt
\[
\infer{\XXi\vdash G_1\vee G_2}
      {\XXi'\vdash G_i\qquad \orExpert{\XXi}{\XXi'}{i}}
\qquad
\infer{\XXi\vdash \exists x. G}
      {\XXi'\vdash G[t/x]\qquad \someExpert{\XXi}{\XXi'}{t}}
\]
\vskip -6pt
\[
\infer{\XXi\vdash t = t}
      {\eqExpert{\XXi}}
\qquad
\infer{\XXi\vdash A}
      {\XXi'\vdash G \quad (A~\hbox{\tt :-}~G)\in\instan\Pscr
                     \quad \unfoldExpert{\XXi}{\XXi'}}
\]
\caption{A proof system augmented with proof certificates and
  additional ``expert'' premises.}
\label{fig:augmented}
\end{figure}

\section{Linear logic as the proof theory of model checking}
\label{sec:ll}

A natural choice is linear logic cousin $\mu\mathrm{MALL}$~\cite{Baelde12}, which is associated to  the \emph{Bedwyr} model-checker. %  In fact, the latter
% has already been used for related aims~\cite{HeathM15}.

\subsection{Horn programs as fixed points}
\label{ssec:lpfixed}

\begin{center}
  [Here we introduce fixed points, and re-state the rules in fig~\ref{fig:augmented} with only unfolding rules. We may also want to introduce our first coinductive examples, probably \texttt{coeval} (it's Horn). 
  ]
\end{center}

\subsection{Onwards to linear logic}
\label{ssec:ll}

\begin{center}
  [
  \begin{itemize}
  \item Here we introduce polarization, focusing. I assume we will be
    using only $\mu$F rather than all muMALL; and if so, perhaps we
    can simplify the syntax of the focusing judgment
  \item Not sure we want to put clercks and experts everywhere, if they play a part only during generation and that should stay in the Horn+nabla fragment
  \item Now we can introduce non Horn examples such as simulation and points out that the left unfolding rules do case analysis.
\end{itemize}
  ]
\end{center}

\subsection{Full system}
\label{ssec:fullmumall}

\begin{center}
  [Here we add the rules for (co)induction, init (and cut?). In the end we take the muFlogic from the JAR paper and the ``completeness'' results thereof]
\end{center}

\section{Implementation and case studies}
\label{sec:case}

\begin{center}
  [Here we discuss the ``architecture'', namely using the lightweight FPC for generators and Bedwyr for testing]
\end{center}  
\subsection{Examples}
\label{ssec:exs}

\begin{center}
  [Which esamples? The ones which work]
\end{center}

\subsection{Refinements}
\label{ssec:refs}

\begin{center}
  [Anything else to make this interesting: integration with
  proof-outlines, pre-computation of terms 'a la Tarau, full
  ``post-mortem'' FPC, that is also for the testing phase, to
  elaborate further the results, for examples for explanations etc., wild ideas about generating infinite terms with observation functions as in \emph{Hipster} \cite{EinarsdottirJP18}]
\end{center}

\section{Related work}
\label{sec:rel}

\begin{center}
  [to much details, just for my own sake, only a small portion will make it in the paper]
\end{center}
  While coinductive logic programming,
see~\cite{Luke07} and~\cite{BasoldKL19} for a much more principled and
in depth treatment, may at first seem to fit the bill, the need to
model infinite \emph{behavior} rather than infinite objects, that is (ir)rational terms on the
domain of discourse, has lead us to adopt a much stronger logic (and
associated proof theory) with explicit rules for induction and
coinduction.
\input{long-related}
\section{Conclusion}
\label{sec:conc}

\bibliographystyle{abbrv}
\bibliography{colp,../ppdp19/l}

@@

%  LocalWords:  Coinductive QuickCheck utf inputenc fontenc Herbrand
%  LocalWords:  corecursion coinductive corecursive QuickChick CBV ir
%  LocalWords:  disjuncts PCFL LTS Harrop PBT Coq's coinductively CCS
%  LocalWords:  intuitionistic FPCs Bedwyr Bedwyr's nabla

>>>>

\RequirePackage[utf8]{inputenc}
\RequirePackage[T1]{fontenc}
\documentclass[a4paper]{easychair}
\usepackage{url,xcolor}
\usepackage{listings}
\usepackage{letltxmacro}
\input{listing-macros}  % Use to get better highlighing of code
\lstset{language=lprolog}
\lstset{language=abella}
\title{On the Proof Theory of Property-Based Testing of Coinductive Specifications, or:\\
 PBT to Infinity and beyond}
\authorrunning{Blanco, Miller and Momigliano}
\titlerunning{PBT to Infinity and
   beyond}
\author{Roberto Blanco\inst{1} \and Dale Miller\inst{2} \and Alberto Momigliano\inst{3}}

@@

@@

While PBT originated in the functional programming community, we have
given in a previous paper (\cite{Blanco0M19}) a reconstruction of some
of its features (operational semantics, different flavors of
generation, shrinking) in purely proof-theoretic terms employing the
framework of \emph{Foundational Proof Certificates}~\cite{chihani17jar}: the
latter, in its full generality, defines a range of proof structures
used in various theorem provers such as resolution refutations,
Herbrand disjuncts, tableaux, etc.
%
In the context of PBT, the proof theory setup is much simpler.
Consider an attempt to find counter-examples to a conjecture of the
form \(\forall x [(\tau(x)\wedge P(x)) \supset Q(x)]\) where $\tau$ is
a typing predicate and $P$ and $Q$ are two other predicates defined
using Horn clause specifications.
%
By negating this conjecture, we attempt to find a (focused) proof of 
\(\exists x [(\tau(x)\land P(x)) \land \neg Q(x)]\).
%
In the focused proof setting, the \emph{positive phase} (where
test cases are generated) is represented by \(\exists x\) and
\((\tau(x)\land P(x))\). 
%
That phase is followed by the \emph{negative phase} (where conjectured
counter-examples are tested) and is represented by \(\neg Q(x)\).
%
%% turns out to be a search for a proof of \(\exists x [(\tau(x)\land
%%   P(x)) \land \neg Q(x)]\) in a focused sequent calculus for
%% (basically) Horn logic, and pre-condition phase, while proving $Q$ can
%% be relegated to deterministic logic programming-like computation (the
%% \emph{negative} phase), interpreting ``$\neg$'' as
%% Negation-as-failure..
%
FPCs are simple logic programs that guide the search for potential
counter-examples using different generation strategies; they further
capture diverse features such as $\delta$-debugging, fault isolation,
explanation, etc.  Such a range of features can be
programmed as the \emph{clerks and experts} predicates that decorate
the sequent rules used in a FPC proof checking kernel: the kernel
is also able to do a limited amount of proof reconstruction.

As explained in~\cite{Blanco0M19}, the standard PBT setup needs little
more than Horn logic. % specifications.
%
However, when addressing infinite computations, we need richer
specifications.  While coinductive logic programming,
see~\cite{Luke07} and~\cite{BasoldKL19} for a much more principled and
in depth treatment, may at first seem to fit the bill, the need to
model infinite \emph{behavior} rather than infinite objects, that is (ir)rational terms on the
domain of discourse, has lead us to adopt a much stronger logic (and
associated proof theory) with explicit rules for induction and
coinduction.

A natural choice for such a logic is the fixed point logic $\cal
G$~\cite{Gacek2012} and its linear logic cousin $\mu\mathrm{MALL}$~\cite{Baelde12}, which are associated to the \emph{Abella} proof
assistant and the \emph{Bedwyr} model-checker.  In fact, the latter
has already been used for related aims~\cite{HeathM15}.

%and supports reasoning over lambda-tree syntax via the $\nabla$ quantifier.

To make things more concrete, consider the usual rules for CBV
evaluation in the $\lambda$-calculus with constants, but define it
\emph{coinductively}, following see~\cite{2007-Leroy-Grall}: using Bedwyr's
concrete syntax, this is written as: 
\begin{lstlisting}
Define coinductive coeval: tm -> tm -> prop by
 coeval (con C) (con C);
 coeval (fun R) (fun R);
 coeval (app M N) V := 
   exists R W, coeval M (fun R) /\ coeval N W /\ coeval (R W) V.
\end{lstlisting}
Is evaluation still deterministic?  And if not, can we find terms
\lstinline{E}, \lstinline{V1}, and \lstinline{V2} such that
\mbox{\lstinline{coeval E V1 /\\ coeval E V2 /\\ (V1 = V2 -> false)}}?\footnote{In
  this case equality is purely syntactical, since by construction terms
  will be ground when compared, but the logic implements a richer
  notion~\cite{GacekMN11}.} Indeed we can, since a divergent term such
as $\Omega$ co-evaluates to anything.  In fact, co-evaluation is not
even type sound in its generality. Our PBT approach aims to find such
counter-examples.

It can also be used to separate various notion of
equivalences in lambda and process calculi: for example,
separating applicative and ground similarity in PCFL~\cite{PITobtpe},
or analogous standard results in the $\pi$-calculus.
%
While analogous goals have been achieved for labeled transition systems
and for CCS (using, for example, the \emph{Concurrency Workbench}), it
is a remarkable feature of the proof-theoretic account that is easy to
generalizes PBT from a system without bindings (say, CCS) to a system
with bindings (say, the $\pi$-calculus).
%
Such ease is possible since proof theory accommodates the
\emph{$\lambda$-tree syntax} approach to treating bindings~\cite{miller18jar}: this approach includes the $\nabla$
quantifier~\cite{miller05tocl} that appears in both Abella and Bedwyr.

In our current setup, we attempt to find counter-examples, using
Bedwyr to execute both the generation of test cases (controlled by
using specific FPCs~\cite{Blanco0M19}) and the testing phase.
%
Such an implementation of PBT has the advantages of allowing us to
piggyback on Bedwyr's facilities for efficient proof 
search via tabling for (co)inductive predicates.
%
There are a couple of treatments of the negation in the testing phase.
%
One approach to eliminating negation from intuitionistic specifications
can be based on the techniques in~\cite{Momigliano00}.
%
Another approach identifies the proof theory behind model checking as
the linear logic $\mu$MALL~\cite{HeathM19} and in that setting,
negations can be eliminated by using De Morgan duality (and
disunification).

\bibliographystyle{abbrv}
\bibliography{../colp,../../ppdp19/l}

@@

%  LocalWords:  Coinductive QuickCheck utf inputenc fontenc Herbrand
%  LocalWords:  corecursion coinductive corecursive QuickChick CBV ir
%  LocalWords:  disjuncts PCFL LTS Harrop PBT Coq's coinductively CCS
%  LocalWords:  intuitionistic FPCs Bedwyr Bedwyr's
